Verification of concurrent design patterns with data

### 2019-02-19 2018-Arch-pNets/arch-pNets.tex
conclusion : Eric + Ludo

\paragraph{Encoding the coordinators.}
The encoding of a coordinator $C$ is a pLTS with: 1)~an
 initial state and an   
$\init$ transition that initialises all the variables to those defined
by the initial valuation $\val{0}{C}$ of $C$, 2)~an action algebra that matches the actions of the coordinator ports but adds, an additional boolean action parameter, and also, for each exported
variable $x$, a corresponding fresh input variable $?x'$ to allow updates during interactions, 3)~pLTS transitions that reflect the original transitions of
$C$ with $\true$ as parameter and 5)~additional loop transitions marked by
$\false$.
\noteLH{enleve le 3 et integre au 2}
Formally,  $\nopri[C] \bydef{=} \pNetTuple{S, s_0, \goesto[{\nopri[C]}]{}}$, such that
%
\begin{itemize}
\item $s_0 \not\in Q_C$ is fresh and $S = Q_C \cup \{s_0\}$,
  \noteEM{En principe les $vars(s)$ devraient etre disjoints, mais je
    veux bien laisser ca sous le tapis tant qu'on n'implemente pas le codage...}
\item $\vars(s) = V_C \cup \setdef{?x'}{x \in \export[C](p), p \in a, s\goesto{a} }$, for all $s \in Q_C$, and $\vars(s_0) =
  \emptyset$,
%% \item the transition relation $\goesto[{\nopri[C]}]{}$ is defined as
%%   the union of three sets of transitions
%%   (with $\mathit{init} \bydef{=} \bigl(x := \val{0}{C}(x)\bigr)^{x \in V_C}$):
%% %
%%   \begin{multline}
%%     \label{eq:nopri:trans}
%%     {\goesto[{\nopri[C]}]{}} \bydef{=}
%%     \bsetdef{\tuple{s, a(\true), g, u, s'}}{
%%       \tuple{s, a, g, u, s'} \in \goesto{}}
%%     \\    
%%     \cup
%%     \bsetdef{\tuple{s, a(\false), \true, \noop, s}}{
%%       s \in Q_C, a \in P_C}
%%     %% \\
%%     \cup
%%     \bigl\{(s_0, \init, \true, \mathit{init}, q^0_C)\bigr\}
%%     \,.
%%   \end{multline}

@@

\paragraph{Encoding the interaction model}
%% First of all, recall that $\nopri[A, \partition ] \bydef{=}
%% \pNetTuple{(\nopri[C])^{C \in \cC}, \partition, \nopri[\Gamma]}$.
%% Thus, 
The holes in $\nopri[A, \partition]$ are indexed by the elements of
the partition \partition.  For the encoding of our running example, we
take $\partition = \bigl\{\{\PortFail, \PortResume\},\{\PortAsk,
\PortReset\}\bigr\}$.  This corresponds to the intuition that the
dangling ports {\PortFail} and {\PortResume} will be provided by a 
monitored component, whereas {\PortAsk} and {\PortReset} correspond to
the actions provided by the ``environment'' (other components of the
system) that are invoked in case of a persistent failure.  As for
the encoding of the coordinators, in the synchronisation vectors of
\nopri[\Gamma], we will associate Boolean values to the actions
corresponding to these ports.

@@

What have we done.

Discussion:

- what did we lern about comparison of the coordination models ?

- extension to other priority features

- proof of correctness ?

>>>>

\paragraph{Encoding the coordinators.}
The encoding of a coordinator $C$ is a pLTS with: 1)~an
 initial state and an   
$\init$ transition that initialises all the variables to those defined
by the initial valuation $\val{0}{C}$ of $C$, 2)~an action algebra that matches the actions of the coordinator ports but adds, an additional boolean action parameter, and also, for each exported
variable $x$, a corresponding fresh input variable $?x'$ to allow updates during interactions, 3)~pLTS transitions that reflect the original transitions of
$C$ with $\true$ as parameter and 4)~additional loop transitions marked by
$\false$.
\noteLH{enleve le 3 et integre au 2}
Formally,  $\nopri[C] \bydef{=} \pNetTuple{S, s_0, \goesto[{\nopri[C]}]{}}$, such that
%
\begin{itemize}
\item $s_0 \not\in Q_C$ is fresh and $S = Q_C \cup \{s_0\}$,
  \noteEM{En principe les $vars(s)$ devraient etre disjoints, mais je
    veux bien laisser ca sous le tapis tant qu'on n'implemente pas le codage...}
\item $\vars(s) = V_C \cup \setdef{?x'}{x \in \export[C](p), p \in a, s\goesto{a} }$, for all $s \in Q_C$, and $\vars(s_0) =
  \emptyset$,
%% \item the transition relation $\goesto[{\nopri[C]}]{}$ is defined as
%%   the union of three sets of transitions
%%   (with $\mathit{init} \bydef{=} \bigl(x := \val{0}{C}(x)\bigr)^{x \in V_C}$):
%% %
%%   \begin{multline}
%%     \label{eq:nopri:trans}
%%     {\goesto[{\nopri[C]}]{}} \bydef{=}
%%     \bsetdef{\tuple{s, a(\true), g, u, s'}}{
%%       \tuple{s, a, g, u, s'} \in \goesto{}}
%%     \\    
%%     \cup
%%     \bsetdef{\tuple{s, a(\false), \true, \noop, s}}{
%%       s \in Q_C, a \in P_C}
%%     %% \\
%%     \cup
%%     \bigl\{(s_0, \init, \true, \mathit{init}, q^0_C)\bigr\}
%%     \,.
%%   \end{multline}

@@

\paragraph{Encoding the interaction model}
%% First of all, recall that $\nopri[A, \partition ] \bydef{=}
%% \pNetTuple{(\nopri[C])^{C \in \cC}, \partition, \nopri[\Gamma]}$.
%% Thus, 
The holes in $\nopri[A, \partition]$ are indexed by the elements of
the partition \partition.  For the encoding of our running example, we
take $\partition = \bigl\{\{\PortFail, \PortResume\},\{\PortAsk,
\PortReset\}\bigr\}$.  This corresponds to the intuition that the
dangling ports {\PortFail} and {\PortResume} will be provided by a \noteLH{monitored ou monitoring???}
monitored component, whereas {\PortAsk} and {\PortReset} correspond to
the actions provided by the ``environment'' (other components of the
system) that are invoked in case of a persistent failure.  As for
the encoding of the coordinators, in the synchronisation vectors of
\nopri[\Gamma], we will associate Boolean values to the actions
corresponding to these ports.

@@

BIP architectures are composition tools that enforce safety properties; the composition of architectures entails the composition of the associated properties.
We have extended architectures with data-sensitive interactions, and proved that this extension still guarantees the preservation of safety properties by architecture composition, under reasonable assumptions. This extends the original compositional methodology offered by BIP architectures. Then we use pNets as a semantic formalism to encode architectures with data. pNet is a low level coordination model for open systems, in which composition preserves bisimulation equivalences. pNet is equipped with tools computing its behavioural semantics in terms of symbolic automata, allowing model-checking and equivalence checking with algorithms relying on SMT engines.
As a result, we obtain automatic and compositional guarantees of safety properties with data where compositionality is given by the BIP architectures, and pNet tools provide automatic verification of the properties of each architecture.
%
%The automatic checking of properties given by the pNet framework
%
%Discussion…:
%-	What did we learn about comparison of the coordination models ? Worth mentioning architectures are flat, so we don’t use the power of hierschical pNets (as providinh=g encapsulation mechanism) ?
%-	What about liveness, fairness ?
%-	Extension to other kind of priorities ?
%-	Shall we make a full proof of correctness of the encoding?
%-	Tooling = implementation of the encoding from BIP arch editing tools + tracing back verification diagnoses to the BIP level.
%
%
%What have we done.
%
%Discussion:
%
%- what did we lern about comparison of the coordination models ?
%
%- extension to other priority features
%
%- proof of correctness ?
